---
title: "Simulation"
author: "Connor Demorest"
date: "5/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

masters = read_csv("data/PicksOddscsv/12Masters.csv")

teams = read_csv("mastpdf.csv")
teams[which(teams$Golfer == "Brendon Grace"), ]$Golfer = "Branden Grace"
teams[which(teams$Golfer == "Erik van Rooyen"), ]$Golfer = "Erik Van Rooyen"
teams[which(teams$Golfer == "Sungjae Im"), ]$Golfer = "Sung Jae Im"
teams[which(teams$Golfer == "Brandon Hankins"), ]$Golfer = "Brandon Harkins"
teams[which(teams$Golfer == "Alex Noren"), ]$Golfer = "Alexander Noren"
teams[which(teams$Golfer == "Rory McIIroy"), ]$Golfer = "Rory McIlroy"
teams[which(teams$Golfer == "Tyrell Hatton"), ]$Golfer = "Tyrrell Hatton"
teams[which(teams$Golfer == "Kyoung-Hoon Lee"), ]$Golfer = "K.H. Lee"
masters[which(masters$Golfer == "Brendon Grace"), ]$Golfer = "Branden Grace"
masters[which(masters$Golfer == "Erik van Rooyen"), ]$Golfer = "Erik Van Rooyen"
masters[which(masters$Golfer == "Sungjae Im"), ]$Golfer = "Sung Jae Im"
masters[which(masters$Golfer == "Brandon Hankins"), ]$Golfer = "Brandon Harkins"
masters[which(masters$Golfer == "Alex Noren"), ]$Golfer = "Alexander Noren"
masters[which(masters$Golfer == "Rory McIIroy"), ]$Golfer = "Rory McIlroy"
masters[which(masters$Golfer == "Tyrell Hatton"), ]$Golfer = "Tyrrell Hatton"
masters[which(masters$Golfer == "Kyoung-Hoon Lee"), ]$Golfer = "K.H. Lee"
```
## Data
Best attempt for language is to say "golfer" for the professional golfers and "player" for the people choosing the professional golfers
Every golfer that has odds. Golfers that were picked with no odds are assumed to be not playing, probably carried over from last week by forgetting to make picks
Masters was chosen because most of the golfers played in it, and it's the biggest tournament of the year so far. Bad thing is that it's a limited field event. 

```{r}
# # Sanity check: does Rahm win the correct proportion of times
# do_once = function(){ 
# sim = sample(x = masters$Golfer, replace = F, size = nrow(masters), prob = masters$WinOdds_prob)
# return(sim[1] == "Jon Rahm")
# }
# replicate(100, replicate(1000, do_once()) %>% mean) %>% density %>% plot
# abline(v = 0.0716)
# # Should be around 0.0716
# masters[which(masters$Golfer=="Jon Rahm"), "WinOdds_prob"]/sum(masters$WinOdds_prob)
# # Seems like it works
```   

Strategy 1: Choose golfers from each group completely randomly.

```{r}
reps = 5000
money = matrix(nrow = 77, ncol = reps)
rownames(money) = c(unique(teams$Player), "Connor_sim") %>% str_sort()
Points_ord = masters %>% select(Points) %>% arrange(desc(Points)) %>% as.vector
                   
for(i in 1:reps) {
samp = masters %>% 
  select(-c(Count, WinOdds)) %>% 
  filter(!is.na(Group)) %>% 
  group_by(Group) %>% 
  sample_n(size = 1, prob = NA) %>% 
  ungroup

picks = masters %>% 
  select(Golfer, Points, Group, WinOdds_prob) %>%
  anti_join(samp, by = "Golfer") %>% 
  sample_n(size = 1, prob = NA) %>%
  bind_rows(samp) %>% 
  select(Golfer, Points) %>% 
  mutate(Player = "Connor_sim")

scramble = data.frame(Golfer = sample(x = masters$Golfer, 
                                      replace = F, 
                                      size = nrow(masters), 
                                      prob = masters$WinOdds_prob)) %>% 
  cbind(Points_ord)

sim_money = teams %>% 
  bind_rows(picks) %>% 
  select(-Points) %>%
  left_join(scramble, by = "Golfer") %>% 
  mutate(across(Points, ~replace(., is.na(.), 0))) %>% 
  group_by(Player) %>% 
  summarise(Tot_points = sum(Points)) %>% 
  arrange(desc(Tot_points)) %>% 
  mutate(Money_won = c(70, 35, rep(0, 75))) %>% 
  select(-Tot_points) %>% 
  arrange(Player)
money[, i] = sim_money$Money_won
}
# If we play this tournament 1000 times how much money do we make?
 apply(money, 1, mean) %>% sort
```

Strategy 2: Our actual picks

```{r}
money2 = matrix(nrow = 77, ncol = reps)
rownames(money2) = c(unique(teams$Player), "Connor_sim") %>% str_sort()
                   
for(i in 1:reps) {
picks = masters %>% 
  filter(Golfer %in% c("Brian Harman", "Brooks Koepka", "Cameron Davis", "Justin Rose", 
                       "Paul Casey", "Tony Finau", "Xander Schauffele")) %>% 
  select(Golfer, Points) %>% 
  mutate(Player = "Connor_sim")

scramble = data.frame(Golfer = sample(x = masters$Golfer, 
                                      replace = F, 
                                      size = nrow(masters), 
                                      prob = masters$WinOdds_prob)) %>% 
  cbind(Points_ord)

sim_money = teams %>% 
  bind_rows(picks) %>% 
  select(-Points) %>%
  left_join(scramble, by = "Golfer") %>% 
  mutate(across(Points, ~replace(., is.na(.), 0))) %>% 
  group_by(Player) %>% 
  summarise(Tot_points = sum(Points)) %>% 
  arrange(desc(Tot_points)) %>% 
  mutate(Money_won = c(70, 35, rep(0, 75))) %>% 
  select(-Tot_points) %>% 
  arrange(Player)
money2[, i] = sim_money$Money_won
}
# If we play this tournament 1000 times how much money do we make?
m2 = apply(money2, 1, mean) %>% sort
```


Strategy 3: Choose golfers with best odds in each group

```{r}
money3 = matrix(nrow = 77, ncol = reps)
rownames(money3) = c(unique(teams$Player), "Connor_sim") %>% str_sort()
                   
for(i in 1:reps) {
# masters %>%
#   group_by(Group) %>% 
#   slice_max(n = 2, WinOdds_prob)
picks = masters %>% 
  filter(Golfer %in% c("Jon Rahm", "Scottie Scheffler", "Cameron Smith", "Adam Scott", "Si Woo Kim", "Russell Henley", "Justin Thomas")) %>% 
  select(Golfer, Points) %>% 
  mutate(Player = "Connor_sim")

scramble = data.frame(Golfer = sample(x = masters$Golfer, 
                                      replace = F, 
                                      size = nrow(masters), 
                                      prob = masters$WinOdds_prob)) %>% 
  cbind(Points_ord)

sim_money = teams %>% 
  bind_rows(picks) %>% 
  select(-Points) %>%
  left_join(scramble, by = "Golfer") %>% 
  mutate(across(Points, ~replace(., is.na(.), 0))) %>% 
  group_by(Player) %>% 
  summarise(Tot_points = sum(Points)) %>% 
  arrange(desc(Tot_points)) %>% 
  mutate(Money_won = c(70, 35, rep(0, 75))) %>% 
  select(-Tot_points) %>% 
  arrange(Player)
money3[, i] = sim_money$Money_won
}
# If we play this tournament 1000 times how much money do we make?
m3 = apply(money3, 1, mean) %>% sort()
```


Strategy 4: Choose golfers with lowest count in each group only

```{r}
money4 = matrix(nrow = 77, ncol = reps)
rownames(money4) = c(unique(teams$Player), "Connor_sim") %>% str_sort()
                   
for(i in 1:reps) {
# masters %>%
#   group_by(Group) %>%
#   slice_min(n = 1, Count)
picks = masters %>% 
  filter(Golfer %in% c("Bryson DeChambeau", "Webb Simpson", "Kevin Na", "Ryan Palmer", "Lucas Herbert", "Erik Van Rooyen", "Robert MacIntyre")) %>% 
  select(Golfer, Points) %>% 
  mutate(Player = "Connor_sim")

scramble = data.frame(Golfer = sample(x = masters$Golfer, 
                                      replace = F, 
                                      size = nrow(masters), 
                                      prob = masters$WinOdds_prob)) %>% 
  cbind(Points_ord)

sim_money = teams %>% 
  bind_rows(picks) %>% 
  select(-Points) %>%
  left_join(scramble, by = "Golfer") %>% 
  mutate(across(Points, ~replace(., is.na(.), 0))) %>% 
  group_by(Player) %>% 
  summarise(Tot_points = sum(Points)) %>% 
  arrange(desc(Tot_points)) %>% 
  mutate(Money_won = c(70, 35, rep(0, 75))) %>% 
  select(-Tot_points) %>% 
  arrange(Player)
money4[, i] = sim_money$Money_won
}
# If we play this tournament 1000 times how much money do we make?
m4 = apply(money4, 1, mean) %>% sort
```


Strategy 5: Choose golfers with best ratio of prob/(count + 1) + best group 1 golfer for WC

```{r}
money5 = matrix(nrow = 77, ncol = reps)
rownames(money5) = c(unique(teams$Player), "Connor_sim") %>% str_sort()
                   
for(i in 1:reps) {
masters %>%
    mutate(rat = WinOdds_prob/(Count+1)) %>%
  group_by(Group) %>%
  slice_max(n = 2, rat)
picks = masters %>% 
  filter(Golfer %in% c("Rory McIlroy", "Webb Simpson", "Tyrrell Hatton", "Sergio Garcia", "Lucas Herbert", 
                       "Erik Van Rooyen", "Bryson DeChambeau")) %>% 
  select(Golfer, Points) %>% 
  mutate(Player = "Connor_sim")

scramble = data.frame(Golfer = sample(x = masters$Golfer, 
                                      replace = F, 
                                      size = nrow(masters), 
                                      prob = masters$WinOdds_prob)) %>% 
  cbind(Points_ord)

sim_money = teams %>% 
  bind_rows(picks) %>% 
  select(-Points) %>%
  left_join(scramble, by = "Golfer") %>% 
  mutate(across(Points, ~replace(., is.na(.), 0))) %>% 
  group_by(Player) %>% 
  summarise(Tot_points = sum(Points)) %>% 
  arrange(desc(Tot_points)) %>% 
  mutate(Money_won = c(70, 35, rep(0, 75))) %>% 
  select(-Tot_points) %>% 
  arrange(Player)
money5[, i] = sim_money$Money_won
}
# If we play this tournament 1000 times how much money do we make?
m5 = apply(money5, 1, mean) %>% sort
```

Strategy 6: Choose golfers with best ratio above 

```{r}
money6 = matrix(nrow = 77, ncol = reps)
rownames(money6) = c(unique(teams$Player), "Connor_sim") %>% str_sort()
                   
for(i in 1:reps) {
# masters %>% 
#     mutate(rat = WinOdds_prob/(Count+1)) %>% 
#   group_by(Group) %>%
#   slice_max(n = 2, rat)
picks = masters %>% 
  filter(Golfer %in% c("Rory McIlroy", "Webb Simpson", "Tyrrell Hatton", "Sergio Garcia", "Lucas Herbert", 
                       "Erik Van Rooyen", "Robert MacIntyre")) %>% 
  select(Golfer, Points) %>% 
  mutate(Player = "Connor_sim")

scramble = data.frame(Golfer = sample(x = masters$Golfer, 
                                      replace = F, 
                                      size = nrow(masters), 
                                      prob = masters$WinOdds_prob)) %>% 
  cbind(Points_ord)

sim_money = teams %>% 
  bind_rows(picks) %>% 
  select(-Points) %>%
  left_join(scramble, by = "Golfer") %>% 
  mutate(across(Points, ~replace(., is.na(.), 0))) %>% 
  group_by(Player) %>% 
  summarise(Tot_points = sum(Points)) %>% 
  arrange(desc(Tot_points)) %>% 
  mutate(Money_won = c(70, 35, rep(0, 75))) %>% 
  select(-Tot_points) %>% 
  arrange(Player)
money6[, i] = sim_money$Money_won
}
# If we play this tournament 1000 times how much money do we make?
m6 = apply(money6, 1, mean) %>% sort
```


Summary: $ won on average 

```{r}
dollas = data.frame(x = factor(c("Dan Dalenberg", "Random", "Actual Picks", "Best Golfer Available", "Low Count", "Balance", "Balance (WC BGA)", "Ploener", "Michael Lunn"), 
                               levels = c("Dan Dalenberg", "Random", "Actual Picks", "Best Golfer Available", 
                                          "Low Count", "Balance", "Balance (WC BGA)", 
                                          "Ploener", "Michael Lunn")),
                    y = c(0.210, 0.987, 1.428, 1.582, 1.659, 2.310, 1.974, 1.519, 3.696))

ggplot(data = dollas, aes(x = x, y = y)) +
  geom_col(fill = "steelblue") +
  geom_hline(yintercept = 1/77*70 + 1/76*35) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.6)) + 
  labs(y = "Average winnings per tourament ($)",
       x = "Simulation method")

```


